The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
Updated abstract based on Simon suggestion
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@478 4525493e-7705-40b1-a816-d608a930855b

\begin{abstract}
A meta-data plays significant role in a large modern enterprises, research experiments,
digital libraries where it comes from different sources and distributed in a 
variety of forms and digital formats. It is organized and managed by constantly
evolving software using both relational and non-relational data sources. There is
a big demand to enable information discovery from multiple sources.
Here we discuss a new data aggregation system which consume and deliver information 
from different relational and non-relational data sources on a concrete example 
of large scale, distributed system of CMS physics experiment at LHC.
\end{abstract}

@@

Once such conglomerate of data-services starts operating an obvious
question arise: how to find out desired information across multiple data-services
in our distributed environment? Even though individual data-services were designed
to answer specific question about data they serve the ability to relate and search
information among them was human task. The growing amount of information
and desire to make cross-service queries force us to design and develop a new
type of system, the Data Aggregation System.

\section{Data Aggregation System\label{DAS}}
The design of DAS system was based on previous studies of CMS 
data bookkeeping system (DBS) usage patterns \cite{DBS07}. We carefully analyzed user
queries placed to DBS, their patterns, frequencies and latency. As a result of this
we created a data discovery service \cite{DD} which came through several iteration
of UI. At the end the end-user interface was based on DBS Query Language \cite{DBS-QL}
with broad variety of template for presentation layer. A quick adoption
and wide spread of DBS QL give us confidence of choosen approach and QL syntax.

Starting from those ground accomplishments we created DAS system as
scalable and pluggable layer sitting on top of existing data-services
within CMS computing infrastructure. Data availability were deligated to individual
data-services. Therefore, we treat DAS as aggregation and cache service 
with the following characteristics:
\begin{itemize}
\item support flexible and intuitive Query Language (QL) to easy look-up our
data;
\item pluggable interface to existing and up-coming data-service APIs;
\item support heterogeneous software environment and distributed nature of data-services;
\item preserve security policies of individual data-services while provide unique
authentication schema for end-users;
\item be flexible to adjust to dynamic nature of data-service APIs, 
due to continuously changing detector and software environments;
\item be transparent to data-service data formats.
\end{itemize}

%\begin{figure}[htb]
%\centering
%\includegraphics[width=150mm]{DAS_architecture.pdf}
%\caption{
%DAS architecture.
%}
%\label{DAS_arch}
%\end{figure}
\begin{figure}[htb]
\centering
\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
\caption{
DAS architecture diagram. It consists of DAS cache server, mapping and analytics DBs.
Once query was placed to DAS an associated
data-service API was invoked, triggering in DAS cache population. Once data become
available in DAS cache user were able to see the results. This so-called on demand
query retrieval system, where data were retrieved and store into cache from
back-end data service. We use DAS analytics DB to establish cache population
for most common queries by using DAS robots.
}
\label{DAS_cache}
\end{figure}

@@

\subsection{Dealing with data-services}
DAS provided pluggable system to add a new CMS data-service. To do that we
identify appropriate APIs, input and output parameters and keys, data-service
format and mapping between data-service, DAS and other sub-systems. 
This information were stored into DAS mapping DB. For instance,
the Data Bookkeeping System (DBS) and data location system (Phedex) both
provide information about CMS data files. In former case, DBS keeps information between
files and other physics  objects, such as run and luminosity, while Phedex provides 
information about file location, such as site and details about storage elements.
Upon user request to find information about a file, based on forementioned
information we were able to identify appropriate set of
data-service APIs, look-up data, re-format their output to DAS notations and
store results into DAS cache.\footebnote{Even though we supported a data retrieval on demand
basis we also store user queries into DAS analytics DB to identify
a common patterns. Based on this information we organized a pre-fetching
strategies.}
Once data were retreived from data-services an additional step to aggregate
matched objects in DAS cache were applied. For instance in above examples
file information came from DBS and Phedex systems. 
Since both data-services output contained file name as a common
key, their output where merged and final result was stored into DAS cache.
This merged (aggregated) results represented a DAS data object. Each DAS data
object contains a standard DAS header, which identify look-up time,
request url, API, parameters and method as well as response version, expiration
time and checksum of the output.

>>>>

\begin{abstract}
%A meta-data plays significant role in a large modern enterprises, research experiments,
%digital libraries where it comes from different sources and distributed in a 
%variety of forms and digital formats. It is organized and managed by constantly
%evolving software using both relational and non-relational data sources. There is
%a big demand to enable information discovery from multiple sources.
%Here we discuss a new data aggregation system which consume and deliver information 
%from different relational and non-relational data sources on a concrete example 
%of large scale, distributed system of CMS physics experiment at LHC.

Meta-data plays signiﬁcant role in a large modern enterprises, 
research experiments, digital libraries where it comes from different 
sources and distributed in a variety of forms and digital formats. 
It is organized and managed by constantly evolving software using 
both relational and non-relational data sources. There is a big 
demand to enable information discovery from multiple sources.

Here we discuss a new data aggregation system which consumes, 
indexes and delivers information from different relational and 
non-relational data sources to answer cross data source queries 
examining metadata associated to petabytes of 
experimental data. We examine simplicity of keyword-based system 
with precise answers of SQL queries under new system where 
aggregated information being collected from various sources 
allowing end-users to place dynamic queries and trigger information 
retrieval on demand. Such close to real-time system based on studies 
of use cases of the CMS particle physics experiment at the LHC which 
uses a large scale, distributed computing system to motivate the work.

%Here we discuss a new data aggregation system which consumes, 
%indexes and delivers information from different relational and 
%non-relational data sources to answer cross data source queries 
%examining metadata associated to petabytes of 
%experimental data. We follow the use cases of the CMS particle 
%physics experiment at the LHC which uses a large scale, distributed 
%computing system system to motivate the work.

\end{abstract}

@@

Once such conglomerate of data-services starts operating an obvious
question arise: how to find out desired information across multiple data-services
in our distributed environment? Even though individual data-services were designed
to answer specific questions about data they serve, the ability to relate and search
information among them was human task. The growing amount of information
and desire to make cross-service queries force us to design and develop a new
type of system, the Data Aggregation System.

\section{Data Aggregation System\label{DAS}}
The design of DAS system was based on previous studies of CMS 
data bookkeeping system (DBS) usage patterns \cite{DBS, DBS07}. We carefully analyzed user
queries placed to DBS, their patterns, frequencies and latency. As a result of this
we created a data discovery service \cite{DD} which came through several iteration
of UI. At the end the end-user interface was based on DBS Query Language (DBS-QL)\cite{DBS-QL}
with broad variety of template for presentation layer. A quick adoption
and wide spread of DBS-QL give us confidence of choosen approach and QL syntax.

Starting from those ground accomplishments we created DAS system as
scalable and pluggable layer sitting on top of the existing data-services
within CMS computing infrastructure. Data availability were deligated to individual
data-services. Therefore, we treat DAS as aggregation and cache service 
with the following characteristics:
\begin{itemize}
\item support flexible and intuitive Query Language (QL) to easy look-up our
data;
\item pluggable interface to existing and up-coming data-service APIs;
\item support heterogeneous software environment and distributed nature of data-services;
\item preserve security policies of individual data-services while provide unique
authentication schema for end-users;
\item be flexible to adjust to dynamic nature of data-service APIs, 
due to continuously changing detector and software environments;
\item be transparent to data-service data formats.
\end{itemize}

%\begin{figure}[htb]
%\centering
%\includegraphics[width=150mm]{DAS_architecture.pdf}
%\caption{
%DAS architecture.
%}
%\label{DAS_arch}
%\end{figure}
\begin{figure}[htb]
\centering
\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
\caption{
DAS architecture diagram. It consists of DAS cache server, mapping and analytics DBs.
Once query was placed to DAS an associated
data-service API was invoked, triggering DAS cache population. Once data become
available in DAS cache user were able to see the results. This on demand
query retrieval system, where data were retrieved and store into cache from
back-end data service. We use DAS analytics DB to establish cache population
for most common queries by using DAS robots.
}
\label{DAS_cache}
\end{figure}

@@

\subsection{Dealing with data-services}
DAS provided pluggable system to add a new CMS data-service. To do that we
identify appropriate APIs, input and output parameters and keys, data-service
format and mapping between data-service, DAS and other sub-systems. 
This information were stored into DAS mapping DB. For instance,
the Data Bookkeeping System (DBS) and data location system (Phedex) both
provide information about CMS data files. In former case, DBS keeps information between
files and other physics  objects, such as run and luminosity, while Phedex provides 
information about file location, such as site and details about storage elements.
Upon user request to find information about a file, based on forementioned
information we were able to identify appropriate set of
data-service APIs, look-up data, re-format their output to DAS notations and
store results into DAS cache.\footnote{Even though we supported a data retrieval on demand
basis we also store user queries into DAS analytics DB to identify
a common patterns. Based on this information we organized a pre-fetching
strategies.}
Once data were retreived from data-services an additional step to aggregate
matched objects in DAS cache were applied. For instance in above examples
file information came from DBS and Phedex systems. 
Since both data-services output contained file name as a common
key, their output where merged and final result was stored into DAS cache.
This merged (aggregated) results represented a DAS data object. Each DAS data
object contains a standard DAS header, which identify look-up time,
request url, API, parameters and method as well as response version, expiration
time and checksum of the output.
